---
title: "neurips_DB"
author: "Kushin Mukherjee"
date: "2023-06-06"
output: html_document
---
```{r}
library(tidyverse)
library(utils)
library(lme4)
library(lmerTest)
library(data.table)
```

```{r}
human_sketch_df<- read.table('../data/human_sketches_R_df.csv',sep=',',header=TRUE)
snp_df<- read.table('../data/snp_R_df.csv',sep=',',header=TRUE)
```

```{r}


h_df<- human_sketch_df%>%subset(model=='human')
h_df$abstraction_c <- (h_df$abstraction - mean(h_df$abstraction))/sd(h_df$abstraction )

nh_df<- human_sketch_df%>%subset(model!='human')
nh_df$abstraction_c <- (nh_df$abstraction - mean(nh_df$abstraction))/sd(nh_df$abstraction )


nh_df$human_entropy<-rep(h_df$entropy,times = 17)
nh_df$human_accuracy<-rep(h_df$mean_top1_accuracy,times = 17)


snp_h_df<- snp_df%>%subset(model=='human')
snp_h_df$abstraction_c<-(snp_h_df$abstraction - mean(snp_h_df$abstraction))/sd(snp_h_df$abstraction )

snp_nh_df<- snp_df%>%subset(model!='human')
snp_nh_df$abstraction_c<-(snp_nh_df$abstraction - mean(snp_nh_df$abstraction))/sd(snp_nh_df$abstraction )

snp_nh_df$human_auc<-rep(snp_h_df$auc,times = 17)

```

## Humans produce more detailed sketches under less stringent time constraints.

```{r}
stroke_complexity_m<- lmer(num_strokes~abstraction +(1+abstraction|concept),data=h_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))

summary(stroke_complexity_m)
```


## Time constraints on sketch production impact sketch understanding


### recognition accuracy
```{r}

### vision models
acc_v_dd_full<-  lmer(mean_top1_accuracy~ abstraction_c+ model+ (1+abstraction_c|concept), data=nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
acc_v_dd_sub<-  lmer(mean_top1_accuracy~ abstraction_c+ (1+abstraction_c|concept), data=nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))


summary(acc_v_dd_full)
anova(acc_v_dd_full,acc_v_dd_sub)


### humans
acc_v_dd_human<-lmer(mean_top1_accuracy~ abstraction_c+ (1+abstraction_c|concept), data=h_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
   
summary(acc_v_dd_human)


```

### soft label distibution entropy

```{r}

### vision models
H_v_dd_full<-  lmer(entropy~ abstraction_c+ model+ (1+abstraction_c|concept), data=nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
H_v_dd_sub<-  lmer(entropy~ abstraction_c+ (1+abstraction_c|concept), data=nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))


summary(H_v_dd_full)
anova(H_v_dd_full,H_v_dd_sub)


### humans
H_v_dd_human<-lmer(entropy~ abstraction_c+ (1+abstraction_c|concept), data=h_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
   
summary(H_v_dd_human)


```

### semantic neighbor preference
```{r}

### vision models
snp_v_dd_full<-  lmer(auc~ abstraction_c+ model+ (1+abstraction_c|concept), data=snp_nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
snp_v_dd_sub<-  lmer(auc~ abstraction_c+ (1+abstraction_c|concept), data=snp_nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))


summary(snp_v_dd_full)
anova(snp_v_dd_full,snp_v_dd_sub)


### humans
snp_v_dd_human<-lmer(auc~ abstraction_c+ (1+abstraction_c|concept), data=snp_h_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
   
summary(snp_v_dd_human)


```

## Vision models vary in their degree of consistency with human recognition performance and response patterns.

### recognition accuracy alignment

```{r}
nh_df_grouped<-nh_df%>%group_by(concept,abstraction,model)%>%summarize(mean_top1_accuracy=mean(mean_top1_accuracy),
                                                      human_accuracy=mean(human_accuracy),
                                                      entropy=mean(entropy),
                                                      human_entropy=mean(human_entropy))

nh_df_grouped$abstraction_c<-(nh_df_grouped$abstraction-mean(nh_df_grouped$abstraction))/sd(nh_df_grouped$abstraction)




acc_aligment_full<-lmer(human_accuracy~ mean_top1_accuracy*abstraction_c+ model+ (1+mean_top1_accuracy*abstraction_c|concept), data=nh_df_grouped)
acc_aligment_sub<- lmer(human_accuracy~ mean_top1_accuracy*abstraction_c+(1+mean_top1_accuracy*abstraction_c|concept), data=nh_df_grouped)
summary(acc_aligment_full)
anova(acc_aligment_full,acc_aligment_sub)
```

### entropy aligment

```{r}

H_aligment_full<-lmer(human_entropy~ entropy*abstraction_c+ model+ (1+entropy*abstraction_c||concept), data=nh_df_grouped)
H_aligment_sub<- lmer(human_entropy~ entropy*abstraction_c+(1+entropy*abstraction_c||concept), data=nh_df_grouped)
summary(H_aligment_full)
anova(H_aligment_full,H_aligment_sub)
```

### snp alignment
```{r}
snp_aligment_full<-lmer(human_auc~ auc*abstraction_c+ model+ (1+auc*abstraction_c|concept), data=snp_nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
snp_aligment_sub<- lmer(human_auc~ auc*abstraction_c+(1+auc*abstraction_c|concept), data=snp_nh_df,control=lmerControl(optimizer="bobyqa",optCtrl=list(maxfun=20000)))
summary(snp_aligment_full)
anova(snp_aligment_full,snp_aligment_sub)
```




